---
title: "DNA damage repair"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(tidyverse)
library(viridis)
library(broom)
library(kableExtra)
```

DDR scores were imported from a perviously cleaned version of the data from Knijnenburg et al. DDR scores were available for 6440 samples. 

```{r ddr_import}
## precalculated by Knijnenburg et al. 

ddr <- "https://github.com/GerkeLab/TCGAhrd/raw/master/data/intermediateFIles/DDRscores.RData"
load(url(ddr))

```

Ancestry estimates were imported from another github repo. Ancestry estimates from the same tissue type were merged with DDR scores. The final dataset contains 6440 samples. 

```{r ancestry_import}

ancestry <- "https://github.com/GerkeLab/TCGAancestry/raw/master/data/admixture_calls.txt"
dat_ancestry <- read.table(ancestry, sep="\t", header = TRUE)

dat_ancestry <- dat_ancestry %>%
  filter(tissue %in% c("Primary Solid Tumor",
                       "Primary Blood Derived Cancer - Peripheral Blood",
                       "Metastatic")) %>%
  mutate(barcode = case_when(
    tissue == "Primary Solid Tumor" ~ paste0(ID, "-01"),
    tissue == "Primary Blood Derived Cancer - Peripheral Blood" ~ paste0(ID, "-03"),
    tissue == "Metastatic" ~ paste0(ID, "-06")
  ))

rm(ancestry)

```

```{r combine_data}

full <- dat %>% 
  select(patient_id, acronym, rppa_ddr_score) %>% 
  left_join(dat_ancestry %>% select(barcode, POP:AFR),
            by = c("patient_id" = "barcode")) %>% 
  filter(!is.na(rppa_ddr_score))

full <- full[!duplicated(full$patient_id),]

rm(dat_ancestry)

```

***

```{r quick_figure, warning=FALSE, fig.pos="center", fig.cap=c("DDR scores by dominant ancestral populations and stratified by cancer type."), fig.width=12, fig.height=12}

ggplot(full, aes(x = POP, y = rppa_ddr_score, fill = as.factor(POP))) + 
  scale_fill_manual(values = rep("white",length(unique(full$POP)))) +
  geom_jitter(position = position_jitterdodge(jitter.width = .95,
                                              jitter.height = 0,
                                              dodge.width = .95),
              aes(fill = as.factor(POP), col = as.factor(POP)), alpha = 0.5) + 
  geom_boxplot(outlier.shape = NA, alpha = 0) + 
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), 
        panel.background = element_blank(), 
        axis.line = element_line(colour = "black")) + 
  guides(fill=FALSE, colour=FALSE) +
  labs(x = "Ancestral Population", y = "DDR Score") + 
  scale_color_viridis_d() + 
  facet_wrap(~acronym)


```

***

Linear model for DDR score by dominant population (ref = "EUR") and adjusting for cancer type. 

```{r quick_tables}

dom_pop <- lm(rppa_ddr_score ~ relevel(full$POP, ref="EUR") + acronym, data = full)

tableX <- tidy(dom_pop) 

tableX <- tableX[2:5,]
tableX$term <- substr(tableX$term, 31,35)

kable(tableX) %>% 
  kable_styling(bootstrap_options = "striped", full_width = F)

```

***

### Blood ancestry 

The previous figure and model were run using the dominant population from blood derived normal samples, rather than the sample that DDR was based off. Only about 10% of samples vary in dominant population between sample types so changes are minimal. 

```{r blood_ancestry, warning=FALSE, fig.pos="center", fig.cap=c("DDR scores by dominant ancestral populations (in normal) and stratified by cancer type."), fig.width=12, fig.height=12}

ancestry <- "https://github.com/GerkeLab/TCGAancestry/raw/master/data/admixture_calls.txt"
blood_ancestry <- read.table(ancestry, sep="\t", header = TRUE)

blood_ancestry <- blood_ancestry %>%
  filter(tissue == "Blood Derived Normal")

blood_dat <- dat %>% 
  mutate(patient_id = substr(patient_id,1,12)) %>%
  inner_join(blood_ancestry,by=c("patient_id" = "ID"))

ggplot(blood_dat, aes(x = POP, y = rppa_ddr_score, fill = as.factor(POP))) + 
  scale_fill_manual(values = rep("white",length(unique(blood_dat$POP)))) +
  geom_jitter(position = position_jitterdodge(jitter.width = .95,
                                              jitter.height = 0,
                                              dodge.width = .95),
              aes(fill = as.factor(POP), col = as.factor(POP)), alpha = 0.5) + 
  geom_boxplot(outlier.shape = NA, alpha = 0) + 
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), 
        panel.background = element_blank(), 
        axis.line = element_line(colour = "black")) + 
  guides(fill=FALSE, colour=FALSE) +
  labs(x = "Ancestral Population - Blood Normal", y = "DDR Score") + 
  scale_color_viridis_d() + 
  facet_wrap(~acronym)

dom_blood_pop <- lm(rppa_ddr_score ~ relevel(blood_dat$POP, ref="EUR") + acronym, data = blood_dat)

tableX <- tidy(dom_blood_pop) 

tableX <- tableX[2:5,]
tableX$term <- substr(tableX$term, 36,38)

kable(tableX) %>% 
  kable_styling(bootstrap_options = "striped", full_width = F)


# entropy <- "https://github.com/GerkeLab/TCGAancestry/raw/master/data/entropy.txt"
# ent <- read.table(entropy, sep="\t", header = TRUE)
# 
# entropy <- ent %>%
#   filter(tissue == "Blood Derived Normal")
# 
# entropy <- dat %>% 
#   mutate(patient_id = substr(patient_id,1,12)) %>%
#   inner_join(entropy,by=c("patient_id" = "ID"))


```